---
title: "Class 11"
author: "Challana Tea"
format: pdf
---

# Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. So, you processed about
~230 samples and did the normalization on a genome level. Now, you want to find whether
there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3
expression.

> Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes. 

```{r}
# Read data into table
exp_data <- read.table('rs8067378_ENSG00000172057.6.csv')
head(exp_data)

# Sample sizes of each genotype
table(exp_data$geno)
```

Find the median expression value of each genotype.
```{r}
aggregate(exp_data$exp, list(exp_data$geno), median)
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)
p <- ggplot(exp_data) +
  aes(x=geno, y=exp, fill=geno) +
  geom_boxplot(notch=TRUE, outlier.shape=NA)
p + geom_jitter(shape=16, position=position_jitter(0.2), alpha = 0.3)
```
The SNP decreases overall expression of ORMDL3

