---
title: "Class 05: Data Visualization with GGPLOT"
author: "Challana Tea"
format: gfm
pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

### R Plotting

The following utilizes "base" R visualization
```{r}
head(cars, n=10)
plot(cars)

```
ggplot is helpful for more complex plots and visualization, but will first need to be installed through install.packages.

```{r}

library(ggplot2)
ggplot(cars)

```
Three layers to each ggplot:
* data (df)
* aesthetics (aes)
* geometrics (geom_type)

```{r}

ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method='lm', se = FALSE) +
  theme_bw()

```
### Plotting DEG Analysis

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

How many genes? --> 5196
How many upregulated --> 127
```{r}
nrow(genes)
table(genes$State)
sum(genes$State == 'up')
```
```{r}
ggplot(genes) +
  aes(x = Condition1, y = Condition2) +
  geom_point()

p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()

p
```

To change the legend colors and axes titles of p
```{r}
p + scale_colour_manual( values=c("blue","gray","red") ) +
  labs(title = "Gene Expression Changes Upon Drug Treatment",
       x = "Control, no drug", y = "Drug Treatment")
```

### Gapminder

Plot GDP per capita against life expectancy from gapminder dataset
```{r}
library(gapminder)
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha = 0.4)
```
How many countries are in this dataset? --> 142
How many years? --> 12

```{r}
length(unique(gapminder$country))
length(unique(gapminder$year))
range(gapminder$year)
```
Smallest population? Which country?

```{r}
ind <- which.min(gapminder$pop)
ind

gapminder$country[ind]
```
```{r}
gapminder[ind,]
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha = 0.4)

ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.4)
```

Splitting by continent
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.4) +
  facet_wrap(~continent)

```