---
title: "Class 8"
author: "Challana Tea"
format: pdf
---

# Import the dataset

Before we can begin our analysis we first have to download and import the data correctly into our R session. Use read.csv()

```{r}
wisc.df <- read.csv("Wisconsincancer.csv",row.names=1)
head(wisc.df)
```
> Q1. How many samples and variables are in this dataset?

```{r}
nrow(wisc.df)
ncol(wisc.df)
```
> Q2. How many M and B samples are there?

```{r}
sum(wisc.df$diagnosis=="M")
sum(wisc.df$diagnosis=="B")
```
```{r}
table(wisc.df$diagnosis)
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
length(grep("_mean",colnames(wisc.df)))
```
> Q. What features are "mean" values (or have "mean" in their name)

```{r}
(grep("_mean",colnames(wisc.df), value=TRUE))
```

I need to remove the first diagnosis column from the data before doing any analysis. I will store it for later as a factor

```{r}
# We can use -1 here to remove the first column
wisc.data <- wisc.df[,-1]
diagnosis <- as.factor(wisc.df$diagnosis)
```

# 2. Principal Component Analysis

The main PCA function in base R is called `prcom()`.

Before doing anything like PCA, it is important to check if the data need to be scaled before performing PCA. Recall two common reasons for scaling data include:

- The input variables use different units of measurement.
- The input variables have significantly different variances.

Looks like we need to scale by setting `scale=TRUE` in our `prcomp()` function call.

```{r}
round(apply(wisc.data, 2, sd), 2)
```
```{r}
wisc.data <- wisc.df[,-32]
```
```{r}
wisc.data <- wisc.data[,-1]
```


## Time for PCA

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```
>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

44.27

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

PC3

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

PC7

```{r}
biplot(wisc.pr)
```
> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

This plot is incredibly messy and very difficult to make out any clear observations.

```{r}
attributes(wisc.pr)
```

Make a little scree plot
```{r}
pr.var <- wisc.pr$sdev^2
#Proportion of variance
pve <- pr.var/sum(pr.var)

plot(pve, typ="o")
```
PC plot

```{r}
library(ggplot2)
pc <- as.data.frame(wisc.pr$x)

ggplot(pc) +
  aes(x=PC1,y=PC2,col=diagnosis) +
  geom_point()
```
# Hierarchical clustering

We can try clustering the original data with `hclust()` or `kmeans()`. First we will scale the data just like we did for PCA

```{r}
data.scaled <- scale(wisc.data)
head(apply(data.scaled,2,sd))
```

```{r}
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
wisc.hclust
```

Plot the dendogram
```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```
To get a cluster membership vector I will use the `cutree()` function and "cut" into 4 or so grps or clusters.

> Q10. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

h = 19

```{r}
grps <- cutree(wisc.hclust, h=19)
table(grps)
```

I can also use the `table()` to cross tabulate...

```{r}
table(diagnosis)
```
```{r}
table(grps,diagnosis)
```

# Clustering on PCA Results

I can cluster in PC-space and use as many or as few PCs as I want.

To start with I will use 3 PCs, that is I will cluster along PC1, PC2, and PC3.

```{r}
pc.dist <- dist(wisc.pr$x[,1:3])
wisc.pr.hclust <- hclust(pc.dist, method="ward.D2")
plot(wisc.pr.hclust)
```
This looks nicer than the previous clustering result. Find the two major clusters with the `cutree()` function
```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```
```{r}
table(grps, diagnosis)
```

We could calculate accuracy - the proportion of samples correct if we take cluster 1 to represent all M and cluster 2 to represent all B.

```{r}
(179+333)/nrow(wisc.data)
```



# Prediction

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```
> Q16. Which of these new patients should we prioritize for follow up based on your results?

Patient 2 should be prioritized.


