---
title: "Class 10: Candy Project"
author: "Challana Tea"
format: pdf
---

Here we take 538's candy dataset and explore it with methods we have been learning in this class

#Import the data

```{r}
candy <- read.csv("candy-data.csv", row.names=1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

# What is your favorite candy?

> Q3. What is your favorite candy in the dataset and what is it's winpercent value?

```{r}
candy["Skittles original", ]$winpercent
```

> Q4. What is the winpercent value for "Kit Kat"?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for "Tootsie Roll Snack Bars"?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The winpercent row

> Q7. What do you think a zero and one represent for the candy$chocolate column?

It indicates if the candy has the chocolate feature.

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```
> Q9. Is the distribution of winpercent values symmetrical?

The values are somewhat symmetrical, but skewed left.

> Q10. Is the center of the distribution above or below 50%?

```{r}
mean(candy$winpercent)
```
The center of distribution is above 50%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
choc.win <- candy[as.logical(candy$chocolate), "winpercent"]
fruity.win <- candy[as.logical(candy$fruity), "winpercent"]
mean(choc.win)
mean(fruity.win)
```

Chocolate candy is higher ranked than fruity candy

> Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruity.win)
```
The difference is statistically significant

# Overall candy rankings

> Q13. What are the five least liked candy types in this set?

```{r}
ord.ind <- order(candy$winpercent)
head(candy[ord.ind,], 5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
ord.ind <- order(candy$winpercent, decreasing=TRUE)
head(candy[ord.ind,], 5)
```
> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  theme(axis.text.y=element_text(size=4)) +
  geom_col()
```
> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  theme(axis.text.y=element_text(size=4)) +
  geom_col()
```
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "red"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  theme(axis.text.y=element_text(size=4)) +
  geom_col(fill=my_cols)
```
> Q17. What is the worst ranked chocolate candy?

Reese's peanut butter cup

> Q18. What is the best ranked fruity candy?

Nik L Nip

# Taking a look at pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 6)
```
> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's miniatures

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
# Exploring the correlation structure

```{r}
library(corrplot)
```
```{r}
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity & chocolate

> Q23. Similarly, what two variables are most positively correlated?

Chocolate & bar

# Principal Component Analysis

Apply PCA using the `prcom()` function to our candy dataset remembering to set the scale=TRUE argument

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
Now we can plot our main PCA score plot of PC1 vs. PC2

```{r}
plot(pca$x[,1:2])
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (pink), fruity (red), other (black)",
       caption="Data from 538")
```

Use plotly to generate an interactive plot where you can use your mouse to see labels.

```{r}
library(plotly)
# ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Pluribus is picked up along with hard and fruity, which makes sense as the more fruity candies with a higher winpercent are those that typically come as hard candy in packs.

